\documentclass[12pt]{../manual}
%____________________________________________________________________________
%
%	READ BEFORE USING
%____________________________________________________________________________
%
% This file will not compile unless all the necessary figures are in subfolder named
% "figures" in the same directory as this file. Note that the first time you compile
% this file in a new computer, the references to the figures will all be ???. You
% will have to compile at least three times for all the references to be correct.
% Lastly, take care to not change/remove anything from above this notice. There is
% no guarantee this will compile without all of the definitions and imports above.
%____________________________________________________________________________
%
%	TITLE AND TABLE OF CONTENTS
%____________________________________________________________________________
\begin{document}
\makeheader{Lab 3}{Fall 2019} % Change Semester and Year as needed
\begin{center}
\textbf{\huge ECE 230L - LAB 3}\\~\\
\textbf{\large INTRODUCTION TO CIRCUIT SIMULATION USING PSPICE}\\~\\
\rule{6.5in}{0.5mm}\\
\end{center}

\tableofcontents

\listoffigures

\newpage
%____________________________________________________________________________
%
%	BODY
%____________________________________________________________________________
\section{Objectives of this Laboratory}
Circuit simulation is an important tool in the analysis and design of microelectronic circuits. Spice is a general-purpose circuit simulation program in which nonlinear dc, nonlinear transient, and linear ac analyses of electronic circuits are carried out. Circuits may contain resistors, capacitors, inductors, mutual inductors, independent voltage and current sources, four types of dependent sources, lossless and lossy transmission lines, switches, uniformly distributed RC lines, and the five most common semiconductor devices: diodes, BJTs, JFETs, MESFETs, and MOSFETs. The version of Spice used in the ECE Department at Duke is PSpice. The objectives of this laboratory session are to introduce you to the basics of PSpice by learning:
\begin{itemize}
\item How to set-up your PSpice simulation environment,
\item How to represent the circuit elements,
\item How to construct the circuits, and
\item How to simulate the circuits.
\end{itemize}

\section{Setting Up a Circuit Using ORCAD Capture}
\begin{enumerate}
\item Open ORCAD Capture CIS
\item Create a new project by selecting \fbox{File} $\to$ \fbox{New} $\to$ \fbox{Project}
\item New your project `Lab 3'
\item Choose \fbox{Analog or Mixed A/D} under the \fbox{Create a New Project Using} menu
\item Select \fbox{Create a blank project} when prompted
\end{enumerate}
Once the new project has been created, circuit design can begin. Sources, components, ground nodes, and wires can be selected using the \fbox{Place} menu.

% Insert image here

PSpice will be used to model the circuit in Figure \ref{fig:exampleCircuit} and perform DC, AC, and transient analysis on the circuit.

\begin{myfigure}[label=fig:exampleCircuit]{Example Circuit}{Example Circuit}
\centering
\begin{circuitikz}[scale=2]
\ctikzset{resistors/scale=1.5,batteries/scale=1.5,grounds/scale=1.5,diodes/scale=1.5}
\draw
(0,0) 	to[short] 		++(2,0)
		node[ground] {}
		to[short] 		++(2,0)
		to[C, l_=${C = \SI{1}{\nano\farad}}$]			++(0,2)
		to[short]		++(-2,0)
		to[R, l_=${R_1 = \SI{1}{\kilo\ohm}}$]			++(-2,0)
		to[battery, l_=$v_{in}$]		++(0,-2)
(2,0)	to[R, *-*, l_=${R_2 = \SI{1}{\kilo\ohm}}$]		++(0,2)
;\end{circuitikz}
\end{myfigure}

To make the circuit, 
\begin{enumerate}
\item Add a DC Voltage Source by following \fbox{Place} $\to$ \fbox{PSPice Component} $\to$ \fbox{Source} $\to$ \fbox{Voltage Source} $\to$ \fbox{DC}
\end{enumerate}

Add a DC voltage source to the circuit by following Place $\to$ PSpice Component $\to$ Source $\to$ Voltage Sources $\to$ DC. After adding the voltage source to the schematic, use the Place $\to$ PSpice Components $\to$ Passives menu to insert the remaining resistors and capacitors. Use Ctrl-R to rotate the components. Use Place $\to$ Wire to connect the circuit nodes. To change values of circuit elements, double click on the element and adjust the desired properties. Finally, add a ground node to the circuit schematic. Follow Place $\to$ Ground and select 0/SOURCE as your ground node.

\section{DC Analysis in PSpice}
To perform a DC analysis of the circuit, you will create a new simulation profile. To create a new profile select PSpice $\to$ New Simulation Profile. Name the new profile `dc' and press Create. To analyze the example circuit, select `DC Sweep' in the Analysis Type drop down menu and use the following parameters:
\begin{itemize}
\item Sweep variable > Voltage source: V1
\item Sweep Type: Linear
\item Start Value: 0
\item End Value: 10
\item Increment: 0.01
\end{itemize}
Press `Apply' and `OK' to save the profile settings. Begin the simulation by selecting PSpice-> Run. To view the circuit behavior at a particular point, follow Trace-> Add Trace to select different values to plot or use the voltage and current markers indicated in Figure 1. Plot source voltage, V(R2), I(R1), and I(R2). Figure 3 shows the circuit schematic and figure 4 shows the result of DC analysis (top plot: current, bottom plot: voltage).

\section{AC Analysis in PSpice}

\subsection{Trace Expressions in PSpice}

\section{Transient Analysis in PSpice}

\section{Practice Example}

\section{Exploration: Thevenin Equivalent Circuits}
\subsection{Purpose}
\subsection{Introduction}
\subsection{Exercise}
\subsection{Practice Exercise: Thevenin Equivalent Circuit}

\newpage
\phantomsection % needed in order to link from table of contents to here
\section*{Grading Rubric}
\addcontentsline{toc}{section}{Grading Rubric} % adds section*{} to table of contents
\markboth{Grading Rubric}{Grading Rubric}
\vfill % used to center table vertically on page
\begin{table}[ht!]
\caption{ECE 230L Laboratory 3 Grading Rubric}
\centering
\begin{tabular}{l|c} \hline
Criteria & Points Possible \\ \hline \hline
\textbf{DC Analysis}			& \textbf{10} \\
Circuit Diagram 				& 5 \\
Waveforms 						& 5 \\ \hline
\textbf{AC Analysis}			& \textbf{10} \\
Circuit Diagram 				& 5 \\
Waveforms 						& 5 \\ \hline
\textbf{Transient Analysis}		& \textbf{10} \\
Circuit Diagram 				& 5 \\ 
Waveforms 						& 5 \\ \hline
\textbf{Practice Exercise}		& \textbf{35} \\
Circuit Diagram 				& 5 \\
DC Analysis						& 10 \\
AC Analysis						& 10 \\
Transient Analysis				& 10 \\ \hline
\textbf{Thevenin Equivalent Example Circuit} & \textbf{20} \\
Circuit Diagram					& 10 \\
$V_{OC}$ and $I_{SC}$ Labeled	& 5 \\
Correct $R_{TH}$ Value			& 5 \\ \hline
\textbf{Thevenin Equivalent Challenge Circuit} & \textbf{15} \\
Circuit Diagram					& 5 \\
$V_{OC}$ and $I_{SC}$ Labeled	& 5 \\
Correct $R_{TH}$ Value			& 5 \\ \hline \hline
Total							& 100 \\ \hline
\end{tabular}
\end{table}
\vfill % used to center table vertically on page
\end{document}
